---
title: "HSB2 Data Analysis"
author: "Jennifer Harrist"
output: html_notebook
---
## Introduction

**The purpose of our research is to investigate the relationship between various variables and student achievement scores in reading, math, and science More specifically, we examine (a) whether a student’s scores in reading and math are related and to what extent this relationship is influenced by socioeconomic status, (b) whether gender impacts a student’s reading and math scores, (c) whether a relationship between race and reading scores exist and to what extent this relationship is influenced by socioeconomic status. Our sample consists of 200 students from different socio-economic backgrounds (low, middle, high) and different ethnicities (White, Hispanic, African-American, Asian).**

We will begin by loading in our data and setting *SES* as a factor.

```{r include = FALSE}

library(tidyverse)
hsb2 <- read_csv("hsb2.csv")
hsb2

SES <- factor(hsb2$ses, 
              levels = c("1", "2", "3"),
              labels = c("low", "middle" ,"high"))
```

## Analysis

Reading is a foundational tool that is required in all content areas. As such, the
instruction of mathematics has changed from being able to compute an algorithm to 
reading and decoding word problems in order to construct the algorithm needed. 
Therefore, It is assumed that a student’s reading ability will be positively related 
to their math scores. 

Hypothesis 1 is proposed: 
H1: Students’ reading score will positively influence to their math scores.

```{r}
p1 <- ggplot(hsb2, aes(x = read, y = math))

p2 <- p1 + geom_point() +
  stat_smooth(method = "lm", fullrange = TRUE, level = 0.70) +
  labs(x = "Reading Scores",
       y = "Math Scores",
       title = "Influence of Math Scores by Reading Scores by SES ",
       caption = "Figure 1: The influence of reading on math scores")
p2
```

In considering H1, a positive relationship exists between reading and math scores, as
seen in Figure 1. Therefore, hypothesis 1 is supported. As reading scores increase,
math scores increase. 

Where reading ability can have a positive influence on math scores, it is assumed 
that differences in SES will influence overall relationship between reading and math 
scores.

```{r}

p3 <- ggplot(hsb2, aes(x = read, y = math, color = SES))
p4 <- p3 + geom_point() +
  stat_smooth(method= "lm", level = 0.70, fullrange = TRUE) + 
  labs(x = "Reading Scores",
       y = "Math Scores",
       title = "Influence of Math Scores by Reading Scores Grouped by SES ",
       caption = "Figure 2: The influence of reading on math scores separated 
       by SES")
p4
```

## Conclusions

A positive relationship for math by reading scores is evident, regardless of student 
SES. Figure 2 indicates that while a positive relationship exists in all SES 
categories, there are different ability levels between the groups, though the difference between them is small. Ability levels are evident between the SES groups, with low
SES having the lowest scores.

